Scalar Feedback based Joint Time-Frequency Precoder Interpolation for MIMO-OFDM Systems

### 2020-03-26 paper/response_letter.tex

\emph{Since the bit budget is small, if the channel varies
sufficiently slowly, the $\theta_{k,l}$ and the $\phi_{k,l}$ parameters
are fed back alternately, while retaining the unsent parameters as
their previous values, as shown in
Fig. 1. This brings down the bit budget to an average of 5 bits for
$\bf{\Theta}$ and 5 bits for $\bf{\Phi}$ over alternate feedback
instances, thereby resulting in an average feedback budget of 5 bits,
as opposed to the 6 bits of feedback used
in [2].}

\textbf{ The so-called time-frequency interpolation method is just a
  straightforward two-dim filter. So I assume the that the more
  important factor for performance improvement would be the use of an
  efficient parameterization for the precoder matrix, which comes from
  ref. [8]. I would expect this paper shed more light on why this
  particular parameterization is more efficient in characterizing
  precoder matrices than the description based on manifolds.  Would
  the authors make more explanations/comments on this?}

\textbf{Response:}

Thank you for the feedback.

It has been recently shown that for asymmetric MIMO channels ($N_t\neq N_R$), precoders that achieve channel capacity via waterfilling reside on non-trivial manifolds like Stiefel and Flag manifolds (Cite Pitavals papers here), which are harder to analyze and characterize than the traditionally considered Grassmannian and Unitary manifolds. A careful study of the manifold methods presented to interpolate and predictively quantize reveals that they also can be interpreted as straightforward 2-D filters using manifold operations like lifting/retraction maps. However, most of the times these operations themselves are suboptimal, like [2] uses quasi-geodesics instead of true geodesics. This leads to a performance degradation. However, the interpolation/prediction methods presented in our manuscript doesn't suffer from these mathematical artefacts and hence outperforms the manifold based methods. To stress explicitly, although both the manifold and the Girvans' rotation parametrization method may be equivalent mathematically, the latter offers more tractability in applying algorithms over them, and gives an hint as to why the latter was also chosen in standards to perform quantization instead of the former.

The interpolation approach that we have
proposed is simple and ends up becoming more accurate over time when
compared to the manifold based interpolation that uses complex
interpolation methods in the manifold space. It is also easier to
track the adaptation of the scalar parameters using few bits. As
observed in Fig. 2, the quantization error obtained when using the
parameter based approach taken in this manuscript is much smaller in
that obtained using the conventional codebooks on the Stiefel
manifold, as discussed in [2]. This results in higher accuracy as
compared to manifolds with similar number of bits.

@@

\textbf{There are some typos/errors to be fixed:}

\textbf{1. Work/works got mixed up for many times.}

Thank you for this comment. We have fixed all instances of this error.

\textbf{2. ``The quantization approach described above...'' Actually,
  parameterization is described without presenting the quantizing
  method.}

@@

\textbf{2. The quantization of $\phi_{k,l}$'s and $\theta_{k,l}$'s mentioned in Subsection III.A and the interpolation
process mentioned in Subsection III.C should be explained more clearly.}

\textbf{Response:}
Thanks for pointing this out. The explanation of the quantization
process has now been made more  clear by altering the text in III.A as
follows:

``\emph{If $n$ bits are allocated to quantize a parameter
for the first time, the range of values that each parameter takes is
divided into $2^n$ optimal intervals using Lloyd's
quantization [16].}''

In addition we have clarified the interpolation process by precisely stating
how we are combining temporally separated precoders as follows:

@@

\emph{To construct precoders for rest of the subcarriers, we will use
precoder parameters from both the ``past'' and ``future'' time
instances to do the interpolation (using backtracking to ensure
causality). When using future time instances, the implication is that the precoder
estimation is improved at a later time instant i.e. after receiving
the feedback of the next set of subcarriers. Note that the precoder
used for transmission is causally obtained, while the ``future''
correction of past precoders merely improves the prediction
accuracy.}

\textbf{3. As mentioned in Fig. 1 and Subsection III.C, the future
  precoder parameters are required to conduct the interpolation. How
  can we obtain the future precoder parameters?}

\textbf{Response:} 
We thank the reviewer for raising this subtle point. We divide the problem we consider into two slightly uncoupled aspects. The receiver estimates the channel matrix and obtains the accurate unquantized precoders. First aspect of the problem considers effective quantization of the precoder by quantizing just the difference between the past and the current precoder, as is typically done in delta quantization approaches. For this aspect we don not use future parameters.

Now, the receiver feeds back these quantized estimates back to the transmitter. The transmitter has to estimate the precoder at the non-fed back time-frequency points, for which it utilizes wait and decode principle to get future precoder data to efficiently interpolate over the previous time precoders. The window size used to get future parameters is low enough to be within the channel coherence time, to make this method physically feasible, however, we would like to highlight that this does add an extra delay to the system, at the expense of improved interpolation error.

\emph{To construct the rest of the subcarriers, we will use
precoder parameters from both the ``past'' and ``future'' and time
instances to do the interpolation (using backtracking to ensure
causality). By future time instances it means that the precoder
estimation is improved at a later time instant i.e. after receiving
the feedback of the next set of subcarriers. Note that the precoder
used for transmission is causally obtained, while the ``future''
correction of past precoders merely improves the prediction (Do you mean interpolation here?)
accuracy.}

### 2020-03-27 paper/response_letter.tex

\textbf{ The so-called time-frequency interpolation method is just a
  straightforward two-dim filter. So I assume the that the more
  important factor for performance improvement would be the use of an
  efficient parameterization for the precoder matrix, which comes from
  ref. [8]. I would expect this paper shed more light on why this
  particular parameterization is more efficient in characterizing
  precoder matrices than the description based on manifolds.  Would
  the authors make more explanations/comments on this?}

\textbf{Response:}

Thank you for the feedback.

It has recently been shown that, for asymmetric MIMO channels
($N_t\neq N_R$), precoders that achieve channel capacity via
waterfilling reside on non-trivial manifolds such as the Stiefel and
Flag manifolds (see references [12-13]) that are harder to analyze and
characterize than the traditionally considered Grassmannian and
Unitary manifolds. A careful study of the manifold methods presented
here to interpolate and predictively quantize reveals that they also
can be interpreted as 2-D filters using manifold operations like
lifting/retraction maps. However, most of these operations themselves
are suboptimal. For instance, the approach in [2] uses quasi-geodesics
instead of true geodesics. This leads to a performance
degradation. However, the interpolation/prediction methods presented
in our manuscript does not suffer from these mathematical artifacts,
and hence outperforms the manifold based methods. To stress this
explicitly, although both the manifold and the Givens' rotation
parametrization method may be close (or equivalent) mathematically,
the latter offers more tractability in applying interpolation
algorithms, and gives hints as to why the latter was also chosen in
standards to perform quantization instead of the former.

@@

In addition we have clarified the interpolation process by precisely stating
how we are combining temporally separated precoders as follows:

\emph{To construct precoders for rest of the subcarriers, we will use
precoder parameters from both the ``past'' and ``future'' time
instances to do the interpolation (using backtracking to ensure
causality). When using future time instances, the implication is that the precoder
estimation is improved at a later time instant i.e. after receiving
the feedback of the next set of subcarriers. Note that the precoder
used for transmission is causally obtained, while the ``future''
correction of past precoders merely improves the prediction
accuracy.}

\textbf{3. As mentioned in Fig. 1 and Subsection III.C, the future
  precoder parameters are required to conduct the interpolation. How
  can we obtain the future precoder parameters?}

\textbf{Response:} We thank the reviewer for raising this subtle
point. We divide the problem we consider into two slightly uncoupled
aspects. The receiver estimates the channel matrix and obtains the
accurate unquantized precoders. The first aspect of the problem
considers effective quantization of the precoder by quantizing just
the difference between the past and the current precoder, as is
typically done in delta quantization approaches. For this aspect we
don not use future parameters.

@@

Now, the receiver feeds back these quantized estimates back to the
transmitter. The transmitter has to estimate the precoder at the
non-fed back time-frequency points, for which it utilizes wait and
decode principle to get future precoder data to efficiently
interpolate over the previous time precoders. The window size used to
get future parameters is low enough to be within the channel coherence
time, to make this method physically feasible, however, we would like
to highlight that this does add an extra delay to the system, at the
expense of improved interpolation error.

\emph{To construct the rest of the subcarriers, we will use
precoder parameters from both the ``past'' and ``future'' and time
instances to do the interpolation (using backtracking to ensure
causality). By future time instances it means that the precoder
estimation is improved at a later time instant i.e. after receiving
the feedback of the next set of subcarriers. Note that the precoder
used for transmission is causally obtained, while the ``future''
correction of past precoders merely improves the interpolation
accuracy.}

### 2020-03-27 paper/main.tex
Minor fixes

\setlength{\textfloatsep}{0.7\baselineskip plus 0.2\baselineskip minus 0.5\baselineskip}
\def\imwidth{0.65}
%package list
\IEEEoverridecommandlockouts
\let\labelindent\relax
% \usepackage{breqn}
\usepackage{enumitem}
\usepackage{fleqn}
\usepackage{cite}
\usepackage{graphicx}
\usepackage[varg]{newtxmath}
\graphicspath{ {images/} }
\usepackage{pdfpages}
\usepackage{wrapfig}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{lettrine}
\usepackage{amsmath}
% \usepackage{titlesec}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{float}
\usepackage[font={footnotesize}]{caption}
\usepackage[numbers,sort,square,compress]{natbib}
\usepackage[para]{footmisc}
\usepackage{xcolor}
\setlength{\belowcaptionskip}{-12pt}
\newcommand{\highlight}[1]{%
  \colorbox{red!50}{$\displaystyle#1$}}

@@

\vspace{-1\baselineskip}
\subsection{Joint Time Frequency Interpolation}
\label{interp}
After obtaining the $\beta$ values (using the single bit feedback) at
the transmitter, we can construct the precoder at subcarrier indices
$pk$ or $pk+q$, keeping with the same notation as the previous
section. To construct precoders for rest of the subcarriers, we will use
precoder parameters from both the ``past'' and ``future'' time
instances to do the interpolation (using backtracking to ensure
causality). When using future time instances, the implication is that the precoder
estimation is improved at a later time instant i.e. after receiving
the feedback of the next set of subcarriers. Note that the precoder
used for transmission is causally obtained, while the ``future''
correction of past precoders merely improves the prediction
accuracy. In Fig.~\ref{fig:adpm-fig}, each parameter is interpolated
using the neighbouring subcarriers using bilinear interpolation
(equivalent to linear interpolation in two dimensions). These
interpolated values are then used to reconstruct the precoding
matrices all together by applying the reverse transformation of
breaking the precoder into independent parameters as given in
Equation~\ref{eq:decompose}. Thus, the approach here is able to
exploit the time and frequency correlation jointly to enhance
performance.

@@

Fig.~\ref{fig:error_decay} shows the rate at which error diminishes
with time for the proposed approach and direct quantization on the
Stiefel manifold as described in~\cite{Gupt1905:Predictive}. Since the
adaptation of independent scalar parameters causes the error to decay
rapidly, the error is lower and the convergence is faster, even though
the starting error is higher in the case of the scalar quantization
approach. The independence of the parameters enables fast
convergence. The Stiefel manifold based approach uses nested codebooks
that may be suboptimal when compared to the parameterization using
angles.

>>>>

\setlength{\textfloatsep}{0.7\baselineskip plus 0.2\baselineskip minus 0.5\baselineskip}
\def\imwidth{0.625}
%package list
\IEEEoverridecommandlockouts
\let\labelindent\relax
% \usepackage{breqn}
\usepackage{enumitem}
\usepackage{fleqn}
\usepackage{cite}
\usepackage{graphicx}
\usepackage[varg]{newtxmath}
\graphicspath{ {images/} }
\usepackage{pdfpages}
\usepackage{wrapfig}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{lettrine}
\usepackage{amsmath}
% \usepackage{titlesec}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{float}
\usepackage[font={footnotesize}]{caption}
\usepackage[numbers,sort,square,compress]{natbib}
\usepackage[para]{footmisc}
\usepackage{xcolor}
\setlength{\belowcaptionskip}{-12pt}
\newcommand{\highlight}[1]{%
  \colorbox{red!50}{$\displaystyle#1$}}

@@

\vspace{-1\baselineskip}
\subsection{Joint Time Frequency Interpolation}
\label{interp}
After obtaining the $\beta$ values (using the single bit feedback) at
the transmitter, we can construct the precoder at subcarrier indices
$pk$ or $pk+q$, keeping with the same notation as the previous
section. To construct precoders for rest of the subcarriers, we will use
precoder parameters from both the ``past'' and ``future'' time
instances to do the interpolation (using backtracking to ensure
causality). When using future time instances, the implication is that the precoder
estimation is improved at a later time instant i.e. after receiving
the feedback of the next set of subcarriers. Note that the precoder
used for transmission is causally obtained, while the ``future''
correction of past precoders merely improves the interpolation
accuracy. In Fig.~\ref{fig:adpm-fig}, each parameter is interpolated
using the neighbouring subcarriers using bilinear interpolation
(equivalent to linear interpolation in two dimensions). These
interpolated values are then used to reconstruct the precoding
matrices all together by applying the reverse transformation of
breaking the precoder into independent parameters as given in
Equation~\ref{eq:decompose}. Thus, the approach here is able to
exploit the time and frequency correlation jointly to enhance
performance.

@@

Fig.~\ref{fig:error_decay} shows the error decay rate with time for
the proposed approach and direct quantization on the Stiefel manifold
as described in~\cite{Gupt1905:Predictive}. Since the adaptation of
independent scalar parameters causes the error to decay rapidly, the
error is lower and the convergence is faster, even for larger starting
errors in the scalar quantization approach. The independence of the
parameters enables fast convergence. The Stiefel manifold based
approach with nested codebooks may thus be suboptimal when compared to
the parameterization using angles.
